### 2014-07-31 Studies/2012_Chicago/Report/Chicago_Fire.tex
Chicago: minor update from Werb comments

Wood can be represented by the the following chemical formula, $\C\H_{1.7}\O_{0.74}\N_{0.002}$, with yields of carbon monoxide ($y_{\mathrm{CO}}=0.004 \; {\mathrm{kg_{CO}}/\mathrm{kg_{wood}}}$) and soot ($y_{\mathrm{C}}=0.015 \; {\mathrm{kg_{C}}/\mathrm{kg_{wood}}}$) found from the SFPE Handbook~\cite{SFPE:Tewarson}. A balanced chemical reaction for wood combustion can written as:

@@

To estimate the energy release rate per unit area (kW/m$^2$) of wood, we use work by Babrauskas and Grayson~\cite{babrauskas1990}. Experiments were conducted in a cone calorimeter\footnote{The cone calorimeter is an experimental apparatus used to gather data about ignition time, mass loss, combustion products, and heat release rate among other properties associated with burning small samples of materials.} to determine the 5 minute average of HRRPUA\footnote{HRRPUA is the heat release rate per unit area.} (kW/m$^2$) for several different types of wood over a range of radiant heat fluxes (kW/m$^2$). Babrauskas and Grayson~\cite{babrauskas1990} suggest that the heat release rate per unit area for a variety of woods is approximately 50~kW/m$^2$.

@@

For the simulation in this study, we can estimate a measure of how well the flow field is resolved by using the non-dimensional expression $D^*/\dx$. Here, $D^*$ is the characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g} }
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}   
From the FDS User Guide~\cite{FDS_Users_Guide}, the characteristic fire diameter is related to the characteristic fire size via the
relation $Q^* = (D^*/D)^{5/2}$. Here, $D$ is the physical diameter of the fire as it is specified in the simulation. Following Eq.~(\ref{eq:mesh}) and using 10 cm grid cells, the ratio of the characteristic fire diameter to the cell size ($D^*/\dx$) is 15.5 for the 3.3~MW porch fire and 23 for the 9~MW attic fire. Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 16~\cite{NUREG_1824} and produced results that were judged to be adequate for engineering calculations. The grid resolution used in this model is within typical engineering $D^*/\dx$ values. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm. Figure~(\ref{fig:geom_grid}) shows the front of the structure rendered in Smokeview with the 10 cm mesh.

@@

\begin{table}
\centering
\captionof{table}{Relevant fire model input parameters}\label{tab:mod_param}
\begin{tabular}{lll}
\toprule[1.5pt]
Parameter & Description & Reference  \\
\midrule
 Simulation Time                   &  5 min                                                        & -- \\[.25cm]
 Grid Cell Size                    &  10 cm                                                        & Section~\ref{mesh} \\ [.25cm]
 Ambient Temperature               &  25 $^{\circ}$C                                               & -- \\ [.1cm]
\multirow{4}{*}{Reaction: Wood}    &  Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$                     & \multirow{4}{*}{Section~\ref{fire}} \\
                                   &  CO Yield: 0.004 \; ${\mathrm{kg_{CO}}/\mathrm{kg_{wood}}}$   & \\
                                   &  Soot Yield: 0.015 \; ${\mathrm{kg_{C}}/\mathrm{kg_{wood}}}$  & \\
                                   &  Heat of Combustion: 16,400~kJ/kg                             & \\ [.25cm]
\multirow{2}{*}{Peak HRR}          &  Attic Fire: 9~MW                                             & \multirow{2}{*}{Section~\ref{fire}}\\ 
                                   &  Porch Fire: 3.3~MW                                           & \\ [.25cm]                     
\multirow{3}{*}{Material: Wood}    &  k: 0.12 \; \si{W/(m.K)})                                     & \multirow{3}{*}{Section~\ref{matl}} \\
                                   &  $\rho$: 510 \; kg/m$^3$                                      & \\
                                   &  c$_{p}$: 1.38 \; \si{kJ/(kg.K)}                              & \\[.25cm]
\multirow{3}{*}{Material: Gypsum}  &  k: 0.17 \; \si{W/(m.K)}                                      &  \multirow{3}{*}{Section~\ref{matl}}\\ 
                                   &  $\rho$: 930 \; kg/m$^3$                                      & \\ 
                                   &  c$_{p}$: 1.09 \; \si{kJ/(kg.K)}                              & \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\begin{figure}[!ht]
\includegraphics[width=.675\textwidth]{../Figures/west_50th_baseline_velo_175}
\input{../Scripts/colorbar_std_velo}
\input{../Scripts/colorbar_metric_velo}
\caption{FDS simulated velocity vectors, 15~s after interior door failure.}
\label{fig:velo_175s}
\end{figure}

@@

\section{Hallway Flow Path}
From on-scene reports (Section~\ref{fire_sum}) and the FDS simulation results (Section~\ref{temp}), the interior conditions on the second floor of the structure were initially tenable. From the NIOSH report, while the IC was inside the structure on the second floor, only light haze and a glow around the door between the hallway and enclosed porch was observed. This indicated the presence of fire in the area of the enclosed porch. The simulation results indicate a high hazard area in the attic and the enclosed porch at this time. By the time the E123 captain and firefighter arrived on the second floor (Table~\ref{tab:firesim}) and approached the rear of the structure, the rear porch was fully involved. Temperature conditions greater than 260~$^{\circ}$C (500~$^{\circ}$F) and pressures greater than 5~Pa were present on the porch side of the steel-faced, wood-framed door at the rear of the structure. Exposure to these conditions weakened the structural integrity of the door, which may have lead to a partial failure. Figure~(\ref{fig:chicago_doorfold}) shows the state of the rear door after the fire. Note the large crease in the steel face of the door just above the knob.

@@

A person is susceptible to second-degree burn injuries if exposed to temperatures greater than 55~$^{\circ}$C (130~$^{\circ}$F)~\cite{designation2003c}. While firefighters wear protective gear, gear only offers a finite amount of protection. The polycarbonate in a self contained breathing apparatus begins to soften when the material temperature reaches approximately 140~$^{\circ}$C (284~$^{\circ}$F)~\cite{mensch2011emergency}. Structural fire fighting coats and pants are tested to withstand temperatures of 260~$^{\circ}$C (500~$^{\circ}$F)~\cite{nfpa2013standard}. Prolonged exposure to elevated temperatures can result in a significant amount of heat transferred to the firefighter, putting him/her at risk. Therefore, for the analysis of the simulated exposure temperatures, we set the maximum visualized temperature to indicate all values that are greater than or equal to 260~$^{\circ}$C (500~$^{\circ}$F). This means that temperatures visualized at the maximum range in the displayed model output are at a minimum of 260~$^{\circ}$C (500~$^{\circ}$F); the actual simulated temperatures could in fact be greater.

>>>>

Wood can be represented by the the following chemical formula, $\C\H_{1.7}\O_{0.74}\N_{0.002}$, with yields of carbon monoxide ($y_{\mathrm{CO}}=0.004 \; \mathrm{kg}/\mathrm{kg}$) and soot ($y_{\mathrm{C}}=0.015 \; \mathrm{kg}/\mathrm{kg}$) per kg of fuel consumed found from the SFPE Handbook~\cite{SFPE:Tewarson}. A balanced chemical reaction for wood combustion can written as:

@@

To estimate the energy release rate per unit area (kW/m$^2$) of wood, work by Babrauskas and Grayson~\cite{babrauskas1990} is used. Experiments were conducted in a cone calorimeter\footnote{The cone calorimeter is an experimental apparatus used to gather data about ignition time, mass loss, combustion products, and heat release rate among other properties associated with burning small samples of materials.} to determine the 5 minute average of HRRPUA\footnote{HRRPUA is the heat release rate per unit area.} (kW/m$^2$) for several different types of wood over a range of radiant heat fluxes (kW/m$^2$). Babrauskas and Grayson~\cite{babrauskas1990} suggest that the heat release rate per unit area for a variety of woods is approximately 50~kW/m$^2$.

@@

For the simulation in this study, a measure of how well the flow field is resolved can be estimated by using the non-dimensional expression $D^*/\dx$. Here, $D^*$ is the characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g} }
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}   
From the FDS User Guide~\cite{FDS_Users_Guide}, the characteristic fire diameter is related to the characteristic fire size via the
relation $Q^* = (D^*/D)^{5/2}$. Here, $D$ is the physical diameter of the fire as it is specified in the simulation. Following Eq.~(\ref{eq:mesh}) and using 10 cm grid cells, the ratio of the characteristic fire diameter to the cell size ($D^*/\dx$) is 15.5 for the 3.3~MW porch fire and 23 for the 9~MW attic fire. Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 16~\cite{NUREG_1824} and produced results that were judged to be adequate for engineering calculations. The grid resolution used in this model is within typical engineering $D^*/\dx$ values. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm. Figure~(\ref{fig:geom_grid}) shows the front of the structure rendered in Smokeview with the 10 cm mesh.

@@

\begin{table}
\centering
\captionof{table}{Relevant fire model input parameters}\label{tab:mod_param}
\begin{tabular}{lll}
\toprule[1.5pt]
Parameter & Description & Reference  \\
\midrule
 Simulation Time                   &  5 min                                                        & -- \\[.25cm]
 Grid Cell Size                    &  10 cm                                                        & Section~\ref{mesh} \\ [.25cm]
 Ambient Temperature               &  25 $^{\circ}$C                                               & -- \\ [.1cm]
\multirow{4}{*}{Reaction: Wood}    &  Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$                     & \multirow{4}{*}{Section~\ref{fire}} \\
                                   &  CO Yield: 0.004 \; $\mathrm{kg}/\mathrm{kg}$   & \\
                                   &  Soot Yield: 0.015 \; $\mathrm{kg}/\mathrm{kg}$  & \\
                                   &  Heat of Combustion: 16,400~kJ/kg                             & \\ [.25cm]
\multirow{2}{*}{Peak HRR}          &  Attic Fire: 9~MW                                             & \multirow{2}{*}{Section~\ref{fire}}\\ 
                                   &  Porch Fire: 3.3~MW                                           & \\ [.25cm]                     
\multirow{3}{*}{Material: Wood}    &  k: 0.12 \; \si{W/(m.K)})                                     & \multirow{3}{*}{Section~\ref{matl}} \\
                                   &  $\rho$: 510 \; kg/m$^3$                                      & \\
                                   &  c$_{p}$: 1.38 \; \si{kJ/(kg.K)}                              & \\[.25cm]
\multirow{3}{*}{Material: Gypsum}  &  k: 0.28 \; \si{W/(m.K)}                                      &  \multirow{3}{*}{Section~\ref{matl}}\\ 
                                   &  $\rho$: 810 \; kg/m$^3$                                      & \\ 
                                   &  c$_{p}$: 1.0 \; \si{kJ/(kg.K)}                              & \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\newpage
\begin{landscape}
\centering
\begin{figure}[!ht]
\includegraphics[width=1.1\textwidth]{../Figures/west_50th_baseline_velo_175}
\input{../Scripts/colorbar_std_velo}
\input{../Scripts/colorbar_metric_velo}
\caption{FDS simulated velocity vectors, 15~s after interior door failure.}
\label{fig:velo_175s}
\end{figure}
\end{landscape}
\newpage

@@

\section{Hallway Flow Path}
From on-scene reports (Section~\ref{fire_sum}) and the FDS simulation results (Section~\ref{temp}), the interior conditions on the second floor of the structure were initially tenable. From the NIOSH report, while the IC was inside the structure on the second floor, only light haze and a glow around the door between the hallway and enclosed porch was observed. This indicated the presence of fire in the area of the enclosed porch. The simulation results indicate a high hazard area in the attic and the enclosed porch at this time. By the time the E123 captain and firefighter arrived on the second floor (Table~\ref{tab:firesim}) and approached the rear of the structure, the rear porch was fully involved. Temperature conditions greater than 260~$^{\circ}$C (500~$^{\circ}$F) were present on the porch side of the steel-faced, wood-framed door at the rear of the structure. Exposure to these conditions weakened the structural integrity of the door, which may have lead to a partial failure. Figure~(\ref{fig:chicago_doorfold}) shows the state of the rear door after the fire. Note the large crease in the steel face of the door just above the knob.

@@

A person is susceptible to second-degree burn injuries if exposed to temperatures greater than 55~$^{\circ}$C (130~$^{\circ}$F)~\cite{designation2003c}. While firefighters wear protective gear, gear only offers a finite amount of protection. The polycarbonate in a self contained breathing apparatus begins to soften when the material temperature reaches approximately 140~$^{\circ}$C (284~$^{\circ}$F)~\cite{mensch2011emergency}. Structural fire fighting coats and pants are tested to withstand temperatures of 260~$^{\circ}$C (500~$^{\circ}$F)~\cite{nfpa2013standard}. Prolonged exposure to elevated temperatures can result in a significant amount of heat transferred to the firefighter, putting him/her at risk. Therefore, for the analysis of the simulated exposure temperatures, the maximum visualized temperature is set to indicate all values that are greater than or equal to 260~$^{\circ}$C (500~$^{\circ}$F). This means that temperatures visualized at the maximum range in the displayed model output are at a minimum of 260~$^{\circ}$C (500~$^{\circ}$F); the actual simulated temperatures could in fact be greater.
